The analogue method for precipitation prediction: finding better analogue situations at a sub-daily time step

### 2017-05-12 Moving temporal window - HESS/template.tex
Obled's corrections.

	\begin{equation}
	\label{eq:S1}
	S_{1}=100 \frac {\displaystyle \sum_{i}^{m} \vert \Delta\hat{z}_{i} - \Delta z_{i} \vert}
	{\displaystyle \sum_{i}^{m} max( \vert \Delta\hat{z}_{i} \vert , \vert \Delta z_{i} \vert ) } ,
	\end{equation}
	where $\Delta \hat{z}_{i}$ is the forecast geopotential height difference between the \textit{i}th pair of adjacent points from the grid of the target situation, $\Delta z_{i}$ is the corresponding observed geopotential height difference in the candidate situation, and $m$ is the number of pairs of adjacent points in the grid. The differences are processed separately in both directions. With smaller S1 values, there is greater similarity in the pressure fields.

@@

	Then, $N_{1}$ dates with the lowest values of S1 are considered as analogues to the target day, where $N_{1}$ needs to be calibrated (Sect. \ref{sec:calibration}). Finally, the daily observed precipitation amount of the corresponding dates provide the empirical conditional distribution considered as the probabilistic prediction for the target day. This method is referred to as 2Z.

@@

	The additional predictor is a moisture index composed of the product of the total precipitable water (TPW) with the relative humidity at 850 hPa (RH850) at 12:00 UTC and 24:00 UTC \citep{Bontron2004}. When adding a second level of analogy, $N_{2}$ dates are subsampled from the $N_{1}$ analogues on the atmospheric circulation, resulting in a smaller number of analogue situations. Moreover, when a second level of analogy is added, a higher number of $N_{1}$ analogues is kept on the first level.

@@

	The calibration is performed in successive steps and aims at minimizing the objective function (CRPS, Eq. \ref{eq:CRPS}). Except for the number of analogues, previously calibrated parameters are generally not reassessed. More advanced techniques such as using genetic algorithms \citep{Horton2017} exist but are beyond the scope of the present study.

@@

	The MTW technique is used to find better analogue situations at different hours of the day rather than comparing the predictors at the same fixed hours. The target situation (the day to predict) is still the same as that in the conventional approach: a daily precipitation total between 06:00 and 30:00 UTC characterised by predictors, which in this case are Z1000 and Z500 at 12:00 and 24:00 UTC, respectively. The difference is that candidate situations are considered not only at the same time (12:00 and 24:00 UTC), but also at other hours by allowing a time shift. Therefore, instead of looking for analogues at a 24-h time step, they are sought at the time step matching the predictor temporal resolution (Fig. \ref{fig:principle}). The ERA-20C dataset was used here to test an MTW with 12-h, 6-h, or 3-h time steps, which are referred to as 12-h MTW, 6-h MTW, and 3-hr MTW, respectively. Thus, the candidates are two (12-h MTW), four (6-h MTW), or eight (3-h MTW) times as many  as in the conventional approach even though they are not fully independent.

	The target situations and their corresponding observed precipitation values do not change. The predictions are asessed on the exact same precipitation time series for both the MTW and the conventional approach so that the performance scores can be directly compared. In order to assess the benefit of searching for analogue situations at a sub-daily time step for quantitative precipitation prediction, an appropriate precipitation archive is required. On the basis of the hourly time series (Sect. \ref{sec:data}), 24-h totals were processed at time steps matching the MTW by building moving 24-h totals on the period 1982--2010, for example starting at 0, 3, ..., 21 h UTC instead of 6 h UTC as for the standard archive.

@@

	When searching for analogues in the first level of analogy, such as on the geopotential heights in the 2Z method (Sect. \ref{sec:analog_method}), up to eight times as many candidates were provided with the MTW than before even though they are not fully independent (Sect. \ref{sec:mtw}). Figure \ref{fig:changes_S1_analogs} shows the changes in the distributions, for target dates over the CP, of the analogy criterion (S1) for the $1^{st}$, $15^{th}$, and $30^{th}$ analogue rank due to the introduction of an MTW with a 12-h, 6-h, or 3-h time step. The shapes of the distributions of the conventional approach and the MTWs were found to be similar. However, the values of the analogy criterion were gradually reduced (shifted to the left) with smaller MTW time steps and were therefore better. The circulation analogy was regularly improved with the doubling of the MTW time step. For analogues with higher ranks (e.g. $15^{th}$ or $30^{th}$), the difference between the two distributions was larger than that for the first rank, which means that the improvement increased with the rank of the analogue.

@@

	The impact of the MTW on the analogy criterion has been analysed per precipitation class, for the target day, for 6-h and 3-h MTWs. The results are summarised in Fig. \ref{fig:changes_S1_precip_threshold} by the median reduction of S1 for the days with precipitation between two thresholds. With a reduction in the number of cases per class, the curves were not as smooth as those in the previous analyses. It is nevertheless clear that the improvements were larger for days with higher precipitation for both the 6-h and 3-h MTW.

@@

	These criterion values were not improved by the MTW because the number of candidates ($N_{1}$) was not higher. The use of an MTW did not increase the sample size in this case because the second level of analogy consists in subsampling in the dates provided by the first level. However, this result of a globally null improvement in the RMSE values does not mean that the 2Z-2MI method cannot be improved by the MTW. After selection of the analogue dates for the synoptic circulation, the new candidate dates did not provide better analogues in terms of moisture. However, the selected dates changed in the first level of analogy and in the final selection; thus, the distributions of precipitation values were different, which can impact the performance of the precipitation prediction.

@@

	Reliability diagrams were also created for the VP for both 2Z and 2Z-2MI methods and for the exceedance of all days percentiles 80~\% (P80, 4 mm), 90~\% (P90, 9.5 mm), and 95~\% (P95, 17.5 mm) at the Ulrichen station (Fig. \ref{fig:reliability_diagrams}). These diagrams plot the observed frequency against the predicted probability of a binary event, which in this case is threshold exceedance. For perfectly calibrated predictions, the curve should be along the diagonal. The VP contains only six years (Sect. \ref{sec:data_methods}); therefore, higher variability is present for higher thresholds, and the curves are smoother and closer to the diagonals for the CP, which is longer (not shown). Figure \ref{fig:reliability_diagrams} shows that the MTW improved the reliability for both 2Z and 2Z-2MI methods and for all thresholds because the curves moved towards the diagonal with decreasing MTW time steps. For P80, the 3-h MTW predictions fit very well the observed frequency. When considering higher thresholds, the 2Z-2MI method was generally better than 2Z. The case of the P95 threshold showed a conditional dry bias associated with larger forecast probabilities for the considered AMs; however, this was substantially reduced by the 3-h MTW. This conditional dry bias means that the AM did not predict the event with a probability equal to 1 for 2Z. Obtaining a predicted probability of exceedance of 1 would mean that all analogue dates for a certain target date would be above the threshold. The number of analogue dates being 30 in this case, it is less likely to have 30 values above the threshold when this one increases. This issue has been addressed by \cite{Marty2010} who proposed a bias correction method for the AM. However, as shown in Figure \ref{fig:reliability_diagrams}, the introduction of an MTW, especially with a 3-h time step, improves this mismatch to some degree.

@@

	Both the analogy criteria (Sect. \ref{sec:changes_analogy_circul}) and the performance scores (Sect. \ref{sec:influence_scores}) were improved to a greater extent for days with heavier precipitation (which are related to higher CRPS values). This is likely attributed to the fact that higher precipitation events are a consequence of atmospheric conditions with greater dynamics such as weather disturbances, which have well-marked temporal evolution. Indeed, the position of the driving elements, such as the low-pressure centres and the fronts, change significantly during the course of a day. These situations are less numerous than anticyclonic situations, which makes it less likely to find very good analogues at the same time of the day. We can, therefore, expect to more significantly improve these situations with greater dynamics when introducing an MTW because better matches to the target situation may be found. In contrast, days with low dynamics in atmospheric circulation, such as anticyclonic situations, will not be radically improved by the introduction of the MTW.

	The MTW improved the prediction for days with heavier precipitation. Therefore, it should improve the prediction of extremes owing to selection of better analogue situations and to addition of possible new extreme precipitation values resulting from moving 24 h totals with a certain time shift. However, even though the distribution of analogue precipitation values should move towards the targeted extreme to provide better prediction, the MTW itself does not enable prediction of extreme events that were not yet observed and are therefore not present in the archive. However, the extremes in AMs can be modelled by extrapolation of a truncated exponential or gamma distribution fitted to the analogue values \citep{Obled2002}. Another possible approach is to combine AMs with other methods \citep[e.g.][]{Chardon2017}. From this perspective, the MTW might improve the prediction of extremes because it improves the distribution of precipitation values for days with higher precipitation, on which post-treatment techniques rely. However, this topic is beyond the scope of the present study.

@@

	The second approach was used to obtain data that are closer to the chronology of the actual precipitation by relying on informative proxy variables during the reconstruction procedure. Data from NCEP/NCAR reanalysis 1 were used for this purpose. Precipitable water and relative humidity at 1000 hPa, 925 hPa, or 850 hPa were assessed at the four points surrounding the catchment and at different times of the day. Precipitation time series from the reanalysis were not considered due to the presence of zeros, which may not always match the prediction; this can lead to an undefined temporal repartition. The best proxy variable, precipitable water, was identified through correlation analyses with the 6-h precipitation time series on the period 1982--2007 (on non-zero precipitation values). When the best proxy was selected, its temporal profile was used to disaggregate the daily precipitation time series.

	These two simple methods did not result in valuable outputs (results not shown). Indeed, the performance improvement brought by the MTW was lost due to the poor quality of the reconstructed precipitation archives. The second method which relied on a proxy variable performed slightly better than the proportional distribution method, but not enough to obtain a relevant time series. Another reanalysis dataset with more accurate moisture variables could produce better proxies.

	These attempts to transpose the MTW on the (usually) longest meteorological archive highlight the importance of the actual rainfall chronology. The MTW is beneficial if the precipitation series are close to the observed values. In the absence of a precipitation series with accurate sub-daily chronology, the introduction of an MTW does not improve the precipitation prediction.

@@

	Improvements of the analogy criterion and the performance scores were both found to be higher for MTWs with a smaller time step of 3 h. A 3-h MTW provided eight times more candidate situations, although they were not fully independent. Because the MTW provides additional situations to the pool of possible analogues, it can be considered as an inflation of the archive, which can be interesting in several applications.

	It is desirable for the AM to have available the biggest pool of candidate analogues. By inflating the archive, the MTW contributes to enriching this pool. However, the use of the MTW relies partly on precipitation series at a sub-daily time step, which are available for the last decades. It would then be even more profitable to apply the MTW to the longest archive possible rather than being limited to the recent period. Attempts to reconstruct longer sub-daily precipitation time series based on simple disaggregation methods were not successful, which demonstrated the importance of the quality of the precipitation archive to be used in the MTW technique. The precipitation prediction provided by the MTW improved only when the precipitation chronology was close to the accurate values. More advanced disaggregation methods might provide suitable sub-daily time series. Alternatively, another option could be to combine both the MTW and the classical approach and to look for analogue situations at a sub-daily time step from the 80's and at a daily time step for the antecedent years. Finally, some other predictands might not need sub-daily total values but point observations such as the occurrence of hail or extreme wind gusts, which make them easier to use with the MTW as no disaggregation is needed.

>>>>

	\begin{equation}
	\label{eq:S1}
	S_{1}=100 \frac {\displaystyle \sum_{i}^{m} \vert \Delta\hat{z}_{i} - \Delta z_{i} \vert}
	{\displaystyle \sum_{i}^{m} max( \vert \Delta\hat{z}_{i} \vert , \vert \Delta z_{i} \vert ) } ,
	\end{equation}
	where $\Delta \hat{z}_{i}$ is the geopotential height difference between the \textit{i}th pair of adjacent points from the grid of the target situation, $\Delta z_{i}$ is the corresponding observed geopotential height difference in the candidate situation, and $m$ is the number of pairs of adjacent points in the grid. The differences are processed separately in both directions. With smaller S1 values, there is greater similarity in the pressure fields.

@@

	Then, $N_{1}$ dates with the lowest values of S1 are considered as analogues to the target day, where $N_{1}$ needs to be calibrated (Sect. \ref{sec:calibration}). Finally, the daily observed precipitation amount of the corresponding dates provide the empirical conditional distribution considered as the probabilistic prediction for the target day. This method based on two geopotential height fields is referred to as 2Z.

@@

	The additional predictor is a moisture index (MI) composed of the product of the total precipitable water (TPW) with the relative humidity at 850 hPa (RH850) at 12:00 UTC and 24:00 UTC \citep{Bontron2004}. When adding a second level of analogy, $N_{2}$ dates are subsampled from the $N_{1}$ analogues on the atmospheric circulation, resulting in a smaller number of analogue situations. Moreover, when a second level of analogy is added, a higher number of $N_{1}$ analogues is kept on the first level.

@@

	The calibration is performed in successive steps and aims at minimizing the objective function described below (CRPS, Eq. \ref{eq:CRPS}). Except for the number of analogues, previously calibrated parameters are generally not reassessed. More advanced techniques such as using genetic algorithms \citep{Horton2017} exist but are beyond the scope of the present study.

@@

	The MTW technique is used to find better analogue situations at different hours of the day rather than comparing the predictors at the same fixed hours. The target situation (the day to predict) is still the same as that in the conventional approach: a daily precipitation total between 06:00 and 30:00 UTC characterised by predictors, which in this case are Z1000 and Z500 at 12:00 and 24:00 UTC, respectively. The difference is that candidate situations are considered not only at the same time (12:00 and 24:00 UTC), but also at other hours by allowing a time shift, according to the temporal resolution of the predictor (Fig. \ref{fig:principle}). The ERA-20C dataset was used here to test an MTW with 12-h, 6-h, or 3-h time steps, which are referred to as 12-h MTW, 6-h MTW, and 3-hr MTW, respectively. Thus, the candidates are two (12-h MTW), four (6-h MTW), or eight (3-h MTW) times as many  as in the conventional approach even though they are not fully independent.

	The target situations and their corresponding observed precipitation values do not change. The predictions are asessed on the exact same precipitation time series for both the MTW and the conventional approach so that the performance scores can be directly compared. However, in order to assess the benefit of searching for analogue situations at a sub-daily time step for quantitative precipitation prediction, an appropriate precipitation archive is required. On the basis of the hourly time series (Sect. \ref{sec:data}), 24-h totals were processed at time steps matching the MTW by building moving 24-h totals on the period 1982--2010, for example starting at 0, 3, ..., 21 h UTC instead of 6 h UTC as for the standard archive.

@@

	When searching for analogues in the first level of analogy, such as on the geopotential heights in the 2Z method (Sect. \ref{sec:analog_method}), up to eight times as many candidates were provided with the MTW than before even though they are not fully independent (Sect. \ref{sec:mtw}). Figure \ref{fig:changes_S1_analogs} shows the changes in the distributions, for target dates over the CP, of the analogy criterion (S1) for the $1^{st}$, $15^{th}$, and $30^{th}$ analogue rank due to the introduction of an MTW with a 12-h, 6-h, or 3-h time step respectively. The shapes of the distributions of the conventional approach and the MTWs were found to be similar. However, the values of the analogy criterion were gradually reduced (shifted to the left) with smaller MTW time steps and were therefore better. The circulation analogy was regularly improved with the doubling of the MTW time step. For analogues with higher ranks (e.g. $15^{th}$ or $30^{th}$), the difference between the two distributions was larger than that for the first rank, which means that the improvement increased with the rank of the analogue.

@@

	The impact of the MTW on the analogy criterion has been analysed per precipitation class, for the target day, for 6-h and 3-h MTWs. The results are summarised in Fig. \ref{fig:changes_S1_precip_threshold} by the median reduction of S1 for the days with precipitation between two a priori selected thresholds. With a reduction in the number of cases per class, the curves were not as smooth as those in the previous analyses. It is nevertheless clear that the improvements were larger for days with higher precipitation for both the 6-h and 3-h MTW.

@@

	These criterion values were not improved by the MTW because the available number of candidates ($N_{1}$) was not higher. The use of an MTW did not increase the sample size in this case because the second level of analogy consists in subsampling in the dates provided by the first level. However, this result of a globally null improvement in the RMSE values does not mean that the 2Z-2MI method cannot be improved by the MTW. After selection of the analogue dates for the synoptic circulation, the new candidate dates did not provide better analogues in terms of moisture. However, the selected dates changed in the first level of analogy and in the final selection; thus, the distributions of precipitation values were different, which can impact the performance of the precipitation prediction.

@@

	Reliability diagrams were also created for the VP for both 2Z and 2Z-2MI methods and for the exceedance of all days percentiles 80~\% (P80, 4 mm), 90~\% (P90, 9.5 mm), and 95~\% (P95, 17.5 mm) at the Ulrichen station (Fig. \ref{fig:reliability_diagrams}). These diagrams plot the observed frequency against the predicted probability of a binary event, which in this case is threshold exceedance. For perfectly calibrated predictions, the curve should be along the diagonal. The VP contains only six years (Sect. \ref{sec:data_methods}); therefore, higher variability is present for higher thresholds, and the curves are smoother and closer to the diagonals for the CP, which is longer (not shown). Figure \ref{fig:reliability_diagrams} shows that the MTW improved the reliability for both 2Z and 2Z-2MI methods and for all thresholds because the curves moved towards the diagonal with decreasing MTW time steps. For P80, the 3-h MTW predictions fit very well the observed frequency. When considering higher thresholds, the 2Z-2MI method was generally better than 2Z. The case of the P95 threshold showed a conditional dry bias associated with larger probabilities for the considered AMs; however, this was substantially reduced by the 3-h MTW. This conditional dry bias means that the AM did not predict the event with a probability equal to 1 for 2Z. Obtaining a predicted probability of exceedance of 1 would mean that all analogue dates for a certain target date would be above the threshold. The number of analogue dates being 30 in this case, it is less likely to have 30 values above the threshold when this one increases. This issue has been addressed by \cite{Marty2010} who proposed a bias correction method for the AM. However, as shown in Figure \ref{fig:reliability_diagrams}, the introduction of an MTW, especially with a 3-h time step, already improves this mismatch to some degree.

@@

	Both the analogy criteria (Sect. \ref{sec:changes_analogy_circul}) quantifying the synoptic similarity and the performance scores of precipitation prediction (Sect. \ref{sec:influence_scores}) were improved to a greater extent for days with heavier precipitation (which are related to higher CRPS values). This is likely attributed to the fact that higher precipitation events are a consequence of atmospheric conditions with greater dynamics such as weather disturbances, which have well-marked temporal evolution. Indeed, the position of the driving elements, such as the low-pressure centres and the fronts, change significantly during the course of a day. These situations are less numerous than anticyclonic situations, which makes it less likely to find very good analogues at the same time of the day. We can, therefore, expect to more significantly improve these situations with greater dynamics when introducing an MTW because better matches to the target situation may be found. In contrast, days with low dynamics in atmospheric circulation, such as anticyclonic situations, will not be radically improved by the introduction of the MTW.

	The MTW improved the prediction for days with heavier precipitation. Therefore, it should improve the prediction of extremes owing to selection of better analogue situations and to addition of possible new extreme precipitation values resulting from moving 24 h totals with a certain time shift. However, even though the distribution of analogue precipitation values should move towards the targeted extreme to provide better prediction, the MTW itself does not enable prediction of extreme events that were not yet observed and are therefore not present in the archive. However, the extremes in AMs can be modelled by extrapolation of a truncated exponential or gamma distribution fitted to the analogue values \citep{Obled2002}, and precipitation values can be proposed for non observed low probabilities. Another possible approach is to combine AMs with other methods \citep[e.g.][]{Chardon2017}. From this perspective, the MTW might improve the prediction of extremes because it improves the distribution of precipitation values for days with higher precipitation, on which post-treatment techniques rely. However, this topic is beyond the scope of the present study.

@@

	The second approach was used to obtain data that are closer to the chronology of the actual precipitation by relying on informative proxy variables during the reconstruction procedure. Data from NCEP/NCAR reanalysis 1 were used for this purpose. Precipitable water and relative humidity at 1000 hPa, 925 hPa, or 850 hPa were assessed at the four points surrounding the study area and at different times of the day. Precipitation time series from the reanalysis were not considered as potential proxy due to the presence of numerous zeros, which may not always match the prediction; this can lead to an undefined temporal repartition. The best proxy variable, precipitable water, was identified through correlation analyses with the 6-h precipitation time series on the period 1982--2007 (on non-zero precipitation values). When the best proxy was selected, its temporal profile was used to disaggregate the daily precipitation time series.

	These two simple (and rather crude) methods did not result in valuable outputs (results not shown). Indeed, the performance improvement brought by the MTW was lost due to the poor quality of the reconstructed precipitation archives. The second method which relied on a proxy variable performed slightly better than the proportional distribution method, but not enough to obtain a relevant time series. Another reanalysis dataset with more accurate moisture variables could produce better proxies.

	These attempts to transpose the MTW on the (usually) longest meteorological archive highlight the importance of the actual rainfall chronology. The MTW is beneficial if the reconstructed precipitation series are close to the observed values. In the absence of a precipitation series with accurate sub-daily chronology, the introduction of an MTW does not improve the precipitation prediction.

@@

	Improvements of the analogy criterion and the performance scores were both found to be higher for MTWs with a smaller time step of 3 h. A 3-h MTW provided eight times more candidate situations, although they were not fully independent. Because the MTW provides additional situations to the pool of possible analogues, it can be considered as an inflation of the archive, which can be interesting in several applications. Indeed, the most accurate reanalysis datasets are limited to the satellite era starting in the 80s, and thus provide shorter archives than reanalysis datasets based on conventional data only. When using such an archive, the MTW approach allows enriching the pool of possible analogues. Moreover, using the MTW approach in the context of a transient climate, where the eventual nonstationarity of the link between predictors and precipitation might discard the relevance of analogues from the distant past, allows relying on a more recent and shorter archive rather than a long one.

	It is desirable for the AM to have available the biggest pool of candidate analogues. By inflating the archive, the MTW contributes to enriching this pool. However, the use of the MTW relies partly on precipitation series at a sub-daily time step, which are available only for the last decades. It would then be even more profitable to apply the MTW to the longest archive possible rather than being limited to the recent period. Attempts to reconstruct longer sub-daily precipitation time series based on simple disaggregation methods were not successful, which demonstrated the importance of the quality of the precipitation archive to be used in the MTW technique. The precipitation prediction provided by the MTW improved only when the precipitation chronology was close to the accurate values. More advanced disaggregation methods might provide suitable sub-daily time series. Alternatively, another option could be to combine both the MTW and the classical approach and to look for analogue situations at a sub-daily time step from the 80's and at a daily time step for the antecedent years. Finally, some other predictands might not need sub-daily total values but point observations such as the occurrence of hail or extreme wind gusts, which make them easier to use with the MTW as no disaggregation is needed.
